---
title: "Bayesian modeling and prediction for movies"
Author: "James Sheldon"
Date: "`r format(Sys.Date(), '%A, %B %d, %Y'`""
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(tidyverse)
library(statsr)
library(BAS)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
#load("movies.Rdata")
```



* * *

## Part 1: Data


* * *

## Part 2: Data manipulation

```{r}
movies <- movies %>% mutate("feature_film" = "no", "drama" = "no", "mpaa_rating_R" = "no", "oscar_season" = "no", "summer_season" = "no")
movies$feature_film[movies$title_type == "Feature Film"] <- "yes"
movies$drama[movies$genre == "Drama"] <- "yes"
movies$mpaa_rating_R[movies$mpaa_rating == "R"] <- "yes"
movies$oscar_season[movies$thtr_rel_month %in% c(10, 11, 12)] <- "yes"
movies$summer_season[movies$thtr_rel_month %in% c(5, 6, 7, 8)] <- "yes"
```

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

